Missing links and the topological robustness of food webs

### 2022-09-20 manuscript/ms_C3_v2.Rmd
Update ms_C3_v2.Rmd

  1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occuring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that predicted food webs usually have greater connectance than observed ones. Thus we expect that predicted food webs are more robust than observed ones. This expectation has never, to our knowledge, been tested, nor has the effect size been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links, though can also make false positives. We did this for 12 different food webs from a wide variety of ecosystems.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, and this can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found the food webs to be robust by 0.52 units and 0.04 units for the most connected and the random species extinction scenarios respectively while no effect in the least connected species extinction scenario.
  5)  These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, and so it would be interesting to compare the dynamical stability of observed and predicted food webs, as well as the topological stability that we have focused on.

@@

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity, and can lead to food web collapse [@ullah2018]. This food web collapse is due to the cascades of secondary extinctions in a food web because of the primary loss of species, for example due to habitat destruction and climate change [@pimmHumanImpactsRates2006; @thomasComparativeLossesBritish2004; @thomasExtinctionRiskClimate2004]. Therefore, research focused on cascading secondary extinctions also known as 'community viability analysis' have been performed extensively in the past few decades to quantify how robust are food webs to species extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005]. And it has been shown that the rate of collapse of a food web is dependent on its structure and complexity [@dunne2002network; @dunneCascadingExtinctionsCommunity2009].

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. These studies have shown that the robustness of the food webs increases with food web connectance. Also, the removal of the most connected species cause considerably more secondary extinctions than the random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. Simulation studies like these which investigate the impact of primary extinctions in a food web to quantify robustness based on its topological structure provide an alternate solution to canonical experiments in natural ecosystems which are not possible or very difficult to conduct [@dunneCascadingExtinctionsCommunity2009].

@@

A key assumption of the observed food webs is that they are very well sampled i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and then do not represent all of the feeding links that occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort as compared to others whereas some trophic links remain unobserved because of linkage constraints irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. Previous studies such as @caronAddressingEltonianShortfall and @guptaSimultaneouslyEstimatingFood2022 have shown that the predicted food webs from these models usually have greater connectance than the observed ones. Therefore, one solution to compensate for undersampling is to use a food web model such as the Allometric Diet Breadth Model (ADBM) to predict the missing links, and to then measure the robustness of the predicted food web. ADBM is a mechanistic model constructed using foraging rules based on the body sizes of prey and predator where trophic interactions satisfying those rules would be predicted by the model which are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links, and also may predict some links that could not, in fact occur.

@@

In our study, we investigate the topological robustness of the ADBM predicted food webs and compare it to that of the observed food webs, and quantify the effect of undersampling i.e overestimation of connectance on the robustness of these predicted food webs. We expect that the ADBM predicted food webs to be more robust as compared to the observed food webs, and for the greater robustness to be related to the amount by which the ADBM overestimates connectance. We do this by simulating primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the implementation of simulation of primary extinctions for three different extinction scenarios on 12 food webs predicted by the ADBM from a wide variety of ecosystems and compute the resultant secondary extinctions. We then compute a robustness metric to quantify the robustness of those predicted food webs and compare them against that of the observed food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. We chose this model because the ADBM can likely predict missing links in the predicted food webs because it consistently overestimated connectance in the predicted food webs as shown by our study in @guptaSimultaneouslyEstimatingFood2022. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables used in the model are the energy content of prey, handling times of the predator on prey, space clearance rate i.e. how fast a predator searches space, and prey densities. All of these variables are derived from the allometric scaling relationship using the body sizes of species. More details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in @petchey2008.

@@

The goodness of fit of the ADBM's predictions depends on the interaction types in the food webs with predictions that are more size-structured such as aquatic and herbivory interactions being better predicted when compared to less size-structured ones such as parasitoid and terrestrial herbivory ones [@petchey2008].

@@

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of the three criteria: removal of (i) the most-connected species, (ii) the least-connected species and (iii) randomly chosen species. The most-connected and least-connected criteria are based on the degree (i.e. the total number of links to resources and from consumers) of species. We considered these mentioned criteria because the random extinction scenario takes into account all the theoretically possible extinction sequences of species that can occur in a food web and the extinction of most-connected species and least-connected species takes into account the two opposite extreme scenarios. These extinction scenarios have been widely used in studying species extinctions and collapse of food webs and other networks [@dunne2002network; @soleComplexityFragilityEcological2001; @dunne2004; @dunneCascadingExtinctionsCommunity2009; @albertStatisticalMechanicsComplex2002].

Given a primary removal of species in a food web if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. Once no more secondary extinctions occurred, then another primary extinction was made, of the next appropriate species depending on the extinction scenario. This process was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of species subjected to primary removals that resulted in a set of extinction (i.e. primary removals plus secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, which results in at least 50% of total species loss [@dunne2002network; @dunne2004; @jonssonReliabilityR50Measure2015; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$), the first primary extinction causes a cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

@@

First, we simulated primary species loss in food webs predicted by the ADBM which had the maximum true skill statistics and compared it to primary species loss in observed food webs. Second, to take into account the uncertainty in robustness in the ADBM predicted food webs we simulated primary species loss and thereby computed robustness for all the ADBM predicted food webs corresponding to the 95% prediction interval of the predicted connectance. In the case of the random extinction scenario, we simulated 1000 random extinction sequences in a single ADBM predicted food web.

## Statistical analysis

In the random extinction scenario, we computed robustness $R_{50}$ for all 1000 independent random extinction sequences and calculated the median as a summary statistics to quantify the average robustness of a single food web. To quantify the effect of undersampling i.e. overestimation of connectance we compute the ratio of the difference in normalised robustness between the ADBM predicted food webs and observed food webs to the difference in their normalised connectance, where normalisation was performed by dividing the variables by their maximum possible values (I.e 0.5 for $R_{50}$ and 1 for connectance). However, we did not perform any statistical significance test because we work with simulated food webs and therefore the p-values of these tests are influenced by the number of model simulations [@whiteEcologistsShouldNot2014].

@@

We first present the secondary extinction curves of the ADBM predicted food webs which had the maximum true skill statistics and the observed food webs for 12 food webs under three different extinction scenarios. We then compare the robustness of all the ADBM predicted food webs within the 95% prediction interval against that of the observed food webs to take into account uncertainty in the robustness across food webs predictions. Finally we quantify the effect of overestimation of connectance on the difference in their robustness estimates.

@@

In the most-connected extinction scenario, the cumulative secondary extinction curves started to rise steeply at a lower number of primary species removal in the observed food webs as compared to the ADBM predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d, e, f, g, h, j, l)). However there were higher number of cumulative secondary extinctions occurring in the ADBM predicted food webs when compared to that of the observed food webs at a high number of primary species removal (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith Pond food web, there were no secondary extinctions for any number of primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the Broadstone Stream (taxonomic aggregation) food web the same was true only for the observed food web but in the ADBM predicted food web there was a steep rise in the cumulative secondary extinctions (Fig. \ref{fig:fig_r1} (b)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of the most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_mostconnected_maxTSS.png")
```

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary extinctions in the ADBM predicted and the observed food webs for five (out of 1000) independent random extinction sequences to show qualitative differences between their secondary extinction curves. The secondary extinctions in the ADBM predicted food webs were more abrupt than that in the observed food webs.

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_ra_extlines_maxTSS.png")
```

Compared to the most-connected and random extinction scenarios, there were fewer secondary extinctions in the least-connected extinction scenario and therefore the secondary extinction curves were flat for most of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs, the extinction curves of the ADBM predicted food webs overlapped with the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)}). Contrary to the most-connected scenario, there was a very high number of secondary extinctions occurring at very low number of primary species removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary \\textbf{removals of the least connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_leastconnected_maxTSS.png")
```

@@

In all of the food webs except Small Reef and Broadstone Stream (taxonomic aggregation), the effect size was positive on average in the most-connected extinction scenario (Fig. \ref{fig:fig_r5} (a)) i.e. overestimation of connectance had a positive effect on the robustness. In the random extinction scenario, there was a positive effect of overestimation of connectance on the robustness for Ythan, Small Reef, Mill Stream, Grasslands, Caricaie Lakes, Capinteria, Broom and Benguela Pelagic (Fig. \ref{fig:fig_r5} (b)). However, the effect size varied across the food webs. In the least-connected extinction scenario, the overestimation of connectance had little effect on average on the difference in the robustness (Fig. \ref{fig:fig_r5} (c)).

@@

The primary removal of species revealed that the ADBM predicted food webs are more robust than the observed food webs on average, with variation within the robustness of the ADBM predicted food webs in the 12 food web ecosystems suggesting that undersampling in food webs can lead to differences in the estimates of food web robustness. The food webs are least robust to primary extinction of the most connected species scenario compared to that of least connected and random extinction scenarios on average. A future development would be to understand how undersampling i.e. overestimation of connectance influences the stability of the dynamics of the ADBM predicted food webs against that of the observed food webs, and compare it with our study.

@@

The robustness of the ADBM predicted food webs were higher than that of the observed food webs on average (Fig. \ref{fig:fig_r4}) for all of the 12 food web ecosystems with some exceptions. This can be attributed to the higher connectance of the ADBM predicted food webs as compared to that of the observed food webs because a species in a food web with a higher connectance has on average more trophic links as compared to a food web with a lower connectance (Fig. \ref{fig:fig_r5}). Our study suggests that it is important to consider undersampling in observed food webs when computing their robustness.

The findings from our study are similar to what had been documented for other food web models [@dunneCascadingExtinctionsCommunity2009], in terms of an increase in the robustness when the connectance of a food web is increased. However, contrary to general expectations [@dunneNetworkStructureBiodiversity2002], food web robustness did not always increase with the connectance (Fig. \ref{fig:fig_r5}). The ADBM consistently predicted higher connectance than that of the observed food webs however, in some cases an increase in connectance resulted in lower robustness (Fig. \ref{fig:fig_r5}) in the predicted food web. For example: the Benguela Pelagic and Small Reef food webs were surprisingly less robust to primary extinctions on average in the least-connected extinction scenario (Fig. \ref{fig:fig_r4} (c) and \ref{fig:fig_r5} (c)). We suspect this is because the ADBM is underestimating the proportion of basal species in the predicted food webs when compared to that of the observed food webs (Fig: 6 (a) in @guptaSimultaneouslyEstimatingFood2022). As a result, these low degree basal species are the ones to be removed at an early stage in the deletion sequence thereby resulting in an earlier food web collapse in the ADBM predicted food web as compared to that of the observed food web (Fig: \ref{fig:fig_r3} (a) and (j)). This suggests that the overestimation of connectance by the ADBM resulted in a more robust food web on average but differences in the predicted food web properties such as underestimation of the proportion of basal species and overestimation of the maximum trophic level (Fig: \ref{fig:fig_a1}) counteracted that effect and led to reduced robustness. This suggests that food web properties other than connectance play an important role in determining the robustness of a food web and therefore should be also taken into account.

@@

The ADBM based on its size-based rules predicts the diet of a consumer and hence is able to predict some links in the predicted food webs which do in reality occur but are actually not observed in the observed food webs perhaps because of low sampling effort. 
<!-- This could also be possible because of rare prey-predator interactions even when sampling is extensive [@grayJoiningDotsAutomated2015; @woodwardChapterIndividualBasedFood2010; @jordanoSamplingNetworksEcological2016].  -->
A future study could be to extend our study to use other food webs models based on size-based rules such as @gravelInferringFoodWeb2013a and @vagnonAllometricNicheModel2021 and compare them with our study. We expect similar results because these food webs are also based on body size rules.

As with any food web model, we expect that there are real false positives in the food webs predicted by the ADBM. Firstly this may be because the ADBM uses only body size as a trait. A trait uncorrelated with the body size may be influential in determining the interaction between two species [@guptaSimultaneouslyEstimatingFood2022]. Secondly, the ADBM can only predict diets that are contiguous with respect to the size of the prey. I.e. it cannot predict that the consumer will consume prey of size 1 and 3, and not consume prey of size 2. However, it is important to note that observed diets are not always contiguous when prey are ordered by their size due to some ecological differences in how predator species choose their prey [@caronAddressingEltonianShortfall]. Therefore a future study could extend our study with other food web models which are not based on body size such as @cattinPhylogeneticConstraintsAdaptation2004 and @allesinaGeneralModelFood2008. We expect to have a difference in results based on whether the trophic interactions in the food webs are governed by size-structured rules or not.

>>>>

  1)  Undersampling can lead to missing trophic interactions in recorded food webs, with potential consequences for the perceived functioning and stability of the food webs. Undersampling can be compensated for by using food web models such as the allometric diet breadth model (ADBM) to predict missing links. Simultaneously, models might predict links which cannot occur, i.e., false positives. 
  2) Previous research shows that (i) food web robustness (the inverse of the number of secondary extinctions occuring due to primary extinctions) increases with connectance (the number of realised trophic links divided by the number of possible links), and (ii) that model predicted food webs usually have greater connectance than observed ones. Thus we expect that predicted food webs are more robust than observed ones. This expectation has never, to our knowledge, been tested, nor has the effect size been quantified.
  3) We fill this research gap by comparing the robustness of observed food webs to the robustness of food webs predicted by a model (the ADBM) that can account for missing links, though which does not predict all links correctly. We did this for 12 different food webs from a wide variety of ecosystems. We used three extinction scenarios: random, most connected, and least connected.
  4)  We found, as expected, that the predicted food webs were more robust than the observed food webs, and this can be attributed to the higher connectance of the predicted food webs. On average, for every one unit of increase in connectance, we found the food webs to be robust by 0.52 units and 0.04 units for the most connected and the random species extinction scenarios respectively while no effect in the least connected species extinction scenario.
  5)  These results show that undersampling can lead to large underestimates of food web robustness that can be compensated for by filling in missing links with food web models. Nevertheless, increased connectance may contribute to lower dynamical stability, and so it would be interesting to compare the dynamical stability of observed and predicted food webs, as well as the topological stability that we have focused on.

@@

Anthropogenic changes such as climate change and habitat destruction are a threat to biodiversity, and can lead to food web collapse [@ullah2018]. This food web collapse is due to the cascades of secondary extinctions in a food web because of the primary loss of species, for example due to habitat destruction and climate change [@pimmHumanImpactsRates2006; @thomasComparativeLossesBritish2004; @thomasExtinctionRiskClimate2004]. An example of a secondary extinction is when a consumer goes extinction because its sole resource species goes extinct. Therefore, research focused on cascading secondary extinctions also known as 'community viability analysis' have been performed extensively in the past few decades to quantify how robust are food webs to species extinction [@dunne2002network; @dunneCascadingExtinctionsCommunity2009; @bergUsingSensitivityAnalysis2011; @ebenmanCOMMUNITYVIABILITYANALYSIS2004; @ebenmanUsingCommunityViability2005]. This research revealed that the rate of collapse of a food web is dependent on its structure and complexity [@dunne2002network; @dunneCascadingExtinctionsCommunity2009].

Simulation of primary species loss has been conducted in observed food webs and model food webs from terrestrial and aquatic ecosystems, where robustness was measured in terms of secondary extinctions [@dunne2002network; @dunneCascadingExtinctionsCommunity2009]. Primary species loss is considered to be extinction of a species due to causes external to the food web. In research with models, the specific cause or causes are usually not specified. Overharvesting of a population which causes extinction could be considered a driver of a primary extinction. A secondary extinction is an extinction caused by a previous extinction (either primary or secondary). Numerous studies have used a topological criteria for assigning a secondary extinction to a species, i.e., if an extinction is of the last resource species of a consumer species, then the consumer species suffers secondary extinction because it has no resources.

When there are few primary extinctions, and these cause many secondary extinctions, a food web is said to be fragile or not robust. When there are few secondary extinctions, a food web is said to be robust. Several studies have shown that the robustness of the food webs increases with food web connectance. Also, the removal (primary extinction) of the most connected species cause considerably more secondary extinctions than the random removals of species [@dunneNetworkStructureBiodiversity2002; @soleComplexityFragilityEcological2001]. Simulation studies like these, which investigate the impact of primary extinctions in a food web to quantify robustness based on its topological structure, provide an alternate solution to canonical experiments in natural ecosystems which are not possible or very difficult to conduct [@dunneCascadingExtinctionsCommunity2009].

@@

A key assumption of the observed food webs is that they are very well sampled i.e. all the links that in reality can occur are represented. However, it is known that not all food webs are very well sampled and then do not represent all of the feeding links that can occur [@caronAddressingEltonianShortfall; @patonaiAggregationIncompleteFood2017; @jordanoSamplingNetworksEcological2016]. Some rare trophic links require more sampling effort as compared to others, whereas some trophic links remain unobserved because of linkage constraints irrespective of sufficient sampling effort [@jordanoSamplingNetworksEcological2016]. Previous studies such as @caronAddressingEltonianShortfall and @guptaSimultaneouslyEstimatingFood2022 have shown that the predicted food webs from these models usually have greater connectance than the observed ones. Therefore, one solution to compensate for undersampling is to use a food web model such as the Allometric Diet Breadth Model (ADBM) to predict the missing links, and to then measure the robustness of the predicted food web. The ADBM is a mechanistic model constructed using foraging rules based on the body sizes of prey and predator where trophic interactions satisfying those rules would be predicted by the model which are perhaps not observed because those interactions are rare. However, this solution is not infallible, as it is likely that the food web model might still miss some links, and also may predict some links that could not, in fact occur.

@@

In our study, we investigate the topological robustness of the ADBM predicted food webs and compare it to that of the observed food webs. We expect the ADBM predicted food webs to be more robust as compared to the observed food webs, and for the greater robustness to be related to the amount by which the ADBM overestimates connectance. We do this by simulating primary species loss in 12 food webs predicted from the ADBM to quantify the secondary loss of extinctions. We use three different approaches of primary species removal: (i) most connected species, (ii) random species and (iii) least connected species to understand if the outcome varies among these approaches.

@@

In the upcoming sections, we present a detailed account of the implementation of simulation of primary extinctions for three different extinction scenarios on 12 food webs predicted by the ADBM from a wide variety of ecosystems and compute the resultant secondary extinctions. We then describe a robustness metric of those predicted food webs and how we compare to properties of the food webs.

@@

The allometric diet breadth model (ADBM) is based on optimal foraging theory, specifically the contingency model [@macarthur1966]. We chose this model because we have already used it to predict missing links in the predicted food webs, whereby it consistently overestimated connectance  @guptaSimultaneouslyEstimatingFood2022. The ADBM predicts the set of prey species a consumer should feed upon to maximise its rate of energy intake [@petchey2008]. The foraging variables used in the model are the energy content of prey, handling times of the predator on prey, space clearance rate i.e. how fast a predator searches space, and prey densities. Each of these variables is derived from the allometric scaling relationship using the body sizes of species. More details on the foraging rules defined in the ADBM and ADBM's predictive power across different food webs can be found in @petchey2008.

@@

**OP: reduce the width of the final column of table 1.**
**OP: state how and why these set of 12 food webs were chosen to study by us.**

The goodness of fit of the ADBM's predictions depends on the interaction types in the food webs. Those with size-structured interactions, such as herbivory in aquatic ecosystems interactions being better predicted when compared to less size-structured ones such as parasitoids and terrestrial herbivory ones [@petchey2008].

@@

We implemented the primary species removal method from @dunneCascadingExtinctionsCommunity2009 by sequentially removing species using one of the three criteria: removal of (i) the most-connected species, (ii) the least-connected species and (iii) randomly chosen species. The most-connected and least-connected criteria are based on the degree (i.e. the total number of links to resources and from consumers) of species. We considered these three criteria because the random extinction scenario takes into account all the theoretically possible extinction sequences of species that can occur in a food web and the extinction of most-connected species and least-connected species takes into account the two opposite extreme scenarios. These extinction scenarios have been widely used in studying species extinctions and collapse of food webs and other networks [@dunne2002network; @soleComplexityFragilityEcological2001; @dunne2004; @dunneCascadingExtinctionsCommunity2009; @albertStatisticalMechanicsComplex2002].

Given a primary removal of species in a food web, if any remaining species lost all of their resource species, or any cannibalistic species lost all of their resource species except the cannibalistic links, they are removed from the web and a secondary extinction was recorded. Secondary extinctions may cause further secondary extinctions, which were also checked for and recorded. Once no more secondary extinctions occurred, then another primary extinction was made, of the next appropriate species depending on the extinction scenario. This process was carried out until all the species were extinct from the food web.

The robustness (R) of a food web was defined as the proportion of species subjected to primary removals that resulted in extinction (primary and secondary extinctions) of some specified proportion of the species. In our study, we use $R_{50}$, the number of primary extinctions divided by the total number of species, which results in at least 50% of total species loss [@dunne2002network; @dunne2004; @jonssonReliabilityR50Measure2015; @dunneCascadingExtinctionsCommunity2009]. Therefore, if primary extinctions never cause any secondary extinctions, the food web is maximally robust and ($R_{50} = 0.50$). Whereas in a minimally robust community ($R_{50} = 1/S$), the first primary extinction causes a cascade of secondary extinctions of at least nearly half of the species in the food web (i.e. at least $S/2 - 1$).

@@

First, we simulated primary species loss in food webs predicted by the ADBM which had the maximum true skill statistics and compared it to primary species loss in observed food webs. Second, to take into account the uncertainty in robustness in the ADBM predicted food webs we simulated primary species loss and thereby computed robustness for all the ADBM predicted food webs corresponding to the 95% prediction interval of the predicted connectance. Furthermore, in the case of the random extinction scenario, we simulated 1000 random extinction sequences in a single ADBM predicted food web.

## Analysis

In the random extinction scenario, we computed robustness $R_{50}$ for all 1000 independent random extinction sequences and calculated the median as a summary statistics to quantify the average robustness of a single food web to random extinction. To quantify the effect of undersampling, i.e. overestimation of connectance, we compute the ratio of the difference in normalised robustness between the ADBM predicted food webs and observed food webs to the difference in their normalised connectance, where normalisation was performed by dividing the variables by their maximum possible values (I.e 0.5 for $R_{50}$ and 1 for connectance). However, we did not perform any statistical significance test because we work with simulated food webs and therefore the p-values of these tests are influenced by the number of model simulations [@whiteEcologistsShouldNot2014].

@@

We first present the secondary extinction curves of the ADBM predicted food webs which had the maximum true skill statistics and of the observed food webs for 12 food webs under three different extinction scenarios. We then compare the robustness of all the ADBM predicted food webs within the 95% prediction interval against that of the observed food webs to take into account uncertainty in the robustness across food webs predictions. Finally we quantify the effect of overestimation of connectance on the difference in their robustness estimates.

@@

In the most-connected extinction scenario, the cumulative secondary extinction curves started to rise steeply at a lower number of primary species removal in the observed food webs as compared to the ADBM predicted food webs for nine food webs (Fig. \ref{fig:fig_r1} (a, c, d, e, f, g, h, j, l)). However there were higher number of cumulative secondary extinctions occurring in the ADBM predicted food webs when compared to that of the observed food webs at a high number of primary species removal in five food webs (Fig. \ref{fig:fig_r1} (a, f, g, h, k)). In the Skipwith Pond food web, there were no secondary extinctions for any number of primary removal of species (Fig. \ref{fig:fig_r1} (i)), whereas in the Broadstone Stream (taxonomic aggregation) food web the same was true only for the observed food web but in the ADBM predicted food web there was a steep rise in the cumulative secondary extinctions (Fig. \ref{fig:fig_r1} (b)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r1} Cumulative secondary extinctions of species resulting from the primary removals of the \\textbf{most connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_mostconnected_maxTSS.png")
```

In Fig. \ref{fig:fig_r2}, we present the cumulative secondary extinctions in the ADBM predicted and the observed food webs for five (out of 1000) independent random extinction sequences to show example variation caused by different random primary extinction orders. The secondary extinctions in the ADBM predicted food webs were more abrupt than that in the observed food webs.

**OP: expand and clarify what is meant by abrupt... describe this in a sentence. (Abrupt could mean different things to different people, so clarification would be good.)

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r2} Cumulative secondary extinctions of species resulting from the primary removals of \\textbf{random species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs for five independent random extinction sequences. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_ra_extlines_maxTSS.png")
```

Compared to the most-connected and random extinction scenarios, there were fewer secondary extinctions in the least-connected extinction scenario and therefore the secondary extinction curves were flat for many of the food webs (Fig. \ref{fig:fig_r3}). In some of the food webs, the extinction curves of the ADBM predicted food webs overlapped with the observed food webs (Fig. \ref{fig:fig_r3} (b, c, g, h, i, k, l)}). In five of the food webs, a very high number of secondary extinctions occurred at very low number of primary species removal (Fig. \ref{fig:fig_r3} (a, d, e, f, j)).

```{r, echo=FALSE, fig.align='center', out.width="400px", fig.cap="\\label{fig:fig_r3} Cumulative secondary extinctions of species resulting from the primary removals of the \\textbf{least connected species} in the ADBM predicted food webs corresponding to the maximum TSS and observed food webs. S denotes the number of species in a food web. The cumulative secondary extinctions of species and the number of species removed have been normalised by the number of species."}
knitr::include_graphics("../results/plot_leastconnected_maxTSS.png")
```

@@

**OP: "Box" in the figure caption is stated, but it is not clear where boxes are in the figure.

In all of the food webs except Small Reef and Broadstone Stream (taxonomic aggregation), the effect size of connectance on robustness was positive on average in the most-connected extinction scenario (Fig. \ref{fig:fig_r5} (a)) i.e. overestimation of connectance had a positive effect on the robustness. In the random extinction scenario, there was a positive effect of overestimation of connectance on the robustness for Ythan, Small Reef, Mill Stream, Grasslands, Caricaie Lakes, Capinteria, Broom and Benguela Pelagic (Fig. \ref{fig:fig_r5} (b)). However, the effect size varied across the food webs. In the least-connected extinction scenario, the overestimation of connectance had little effect on average on the difference in the robustness (Fig. \ref{fig:fig_r5} (c)).

**OP: In figure 5c it looks like most of the data is less than zero, i.e. higher connectance is associated with lower robustness.**

@@

As expected, the ADBM predicted food webs were more robust than the observed food webs on average. The considerable variation of the robustness of the ADBM predicted food webs suggests, however, that undersampling in food webs can lead to considerable uncertainty in the estimates of food web robustness, even when a model is used to compensate for undersampling. As was previously found, the food webs are least robust to primary extinction of the most connected species compared to that of least connected and random extinction scenarios on average. A future development would be to understand how undersampling, i.e. overestimation of connectance, influences the stability of the dynamics of the ADBM predicted food webs against that of the observed food webs, and compare it with the patterns in our study in which extinction occur only by topological criteria. **OP: what might one expect, based on previous work/evidence/results?*

@@

As mentioned, the robustness of the ADBM predicted food webs was higher than that of the observed food webs on average (Fig. \ref{fig:fig_r4}) for all of the 12 food web ecosystems (with some exceptions). This is likely due to the higher connectance of the ADBM predicted food webs as compared to that of the observed food webs because a species in a food web with a higher connectance has on average more trophic links as compared to a food web with a lower connectance (Fig. \ref{fig:fig_r5}). Our study suggests that it is important to consider undersampling in observed food webs when computing their robustness.

The findings from our study are similar to what had been documented for other food web models [@dunneCascadingExtinctionsCommunity2009], in terms of an increase in the robustness when the connectance of a food web is increased. However, contrary to general expectations [@dunneNetworkStructureBiodiversity2002], food web robustness did not always increase with the connectance (Fig. \ref{fig:fig_r5}). **OP: the next sentence seems to be repeating the previous two, rather than adding some content. Please clarify what development is provided in the next sentence** The ADBM usually predicted higher connectance than that of the observed food webs however, in some cases an increase in connectance resulted in lower robustness (Fig. \ref{fig:fig_r5}) in the predicted food web. For example: the Benguela Pelagic and Small Reef food webs were surprisingly less robust to primary extinctions on average in the least-connected extinction scenario (Fig. \ref{fig:fig_r4} (c) and \ref{fig:fig_r5} (c)). **OP. next sentence is my suggestion, please check it is accurate** In these two food webs, extinction of the least connected species could cause an almost complete, or complete collapse of the food web. We suspect this is because the ADBM is underestimating the proportion of basal species in the predicted food webs when compared to that of the observed food webs (Fig: 6 (a) in @guptaSimultaneouslyEstimatingFood2022). As a result, these low degree basal species are the ones to be removed at an early stage in the deletion sequence thereby resulting in an earlier food web collapse in the ADBM predicted food web as compared to that of the observed food web (Fig: \ref{fig:fig_r3} (a) and (j)). This suggests that the overestimation of connectance by the ADBM resulted in a more robust food web on average but differences in the predicted food web properties such as underestimation of the proportion of basal species and overestimation of the maximum trophic level (Fig: \ref{fig:fig_a1}) counteracted that effect and led to reduced robustness. This suggests that food web properties other than connectance play an important role in determining the robustness of a food web and therefore should be also taken into account (**OP: add a citation to one or more of the published studies that have shown this.).

@@

**OP: the next sentence does not add any value, which is not so good in a two sentence paragraph. You could delete the first sentence and replace it with some sentences along the lines of: One of the choices we made was to use the ADBM, whereas we could have chosed to use other models. Alternative include modelX, modelY,... And then describe those models in some detail and the continue what remains here.** The ADBM based on its size-based rules predicts the diet of a consumer and hence is able to predict some links in the predicted food webs which do in reality occur but are actually not observed in the observed food webs perhaps because of low sampling effort. 
<!-- This could also be possible because of rare prey-predator interactions even when sampling is extensive [@grayJoiningDotsAutomated2015; @woodwardChapterIndividualBasedFood2010; @jordanoSamplingNetworksEcological2016].  -->
A future study could be to extend our study to use other food webs models based on size-based rules such as @gravelInferringFoodWeb2013a and @vagnonAllometricNicheModel2021 and compare them with our study. We expect, however, similar results because these food webs are also based on body size rules.
